---
title: "Compartment Switching Graphic Choices"
author: Meng Wang
date: "`r format(Sys.time(), '%B %d, %Y')`"
always_allow_html: true
output:
    pdf_document
---

```{r loadLibrary,  include=FALSE, echo=FALSE, messages = FALSE, warnings=FALSE}
library(ggplot2)
library(reshape2)
library(viridis)
library(dplyr)
library(hrbrthemes)
library(gt)
theme_set(theme_bw())
theme_update(
    plot.title = element_text(hjust = 0.5)
)
library(knitr)
library(grid)
library(gridExtra)
get_legend <- function(myggplot) {
    tmp <- ggplot_gtable(ggplot_build(myggplot))
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-aox")
    legend <- tmp$grobs[[leg]]
    return(legend)
}
library(gt)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
```

# Summary:
- Goal: P3F/MYOD/K27ac association with discordant compartment.
- Methods: Distribution of binding sites per category (A-A, A-B, B-A, B-B).
- Problem: Compare per category or per 1000 binding sites?
    - For K27ac and MYOD, FP is quite comparable becauase they have similar number of binding sites in FP.
    - Normalizing by number of bins: P3F is about 1/9 binding sites of MYOD1, so it would sit lower than both (Fig1).
    - Normalize by binding sites: fraction is really low. (Fig 2)
    - Normalize by both: I feel like P3F is overblown. The fact that there are only 2000 or so consensus sites in FP means something.
    I am not sure if downscaling K27ac/MYOD is the right approach. (Fig 4)
- Additional graphical presentation choices: FP/FN as facets (Fig1-3), or Factor as facets (Fig4)?

```{r}

echo "cell factor AB BA AA BB sites" > switch_table_tf.txt


bed="/home/gdstantonlab/mxw010/Data/Ben/pc-chip-seq/GSL-BS-2005/H3K27AC/RH4/align_primary/chip/b6b42932-8417-4ffd-acfe-6c4ce74da28f/call-call_peak/shard-0/execution/SEG0134_S2_L003_R1_001_val_1.srt.nodup_x_SEG0133_S1_L003_R1_001_val_1.srt.nodup.pval0.01.500K.bfilt.narrowPeak.gz"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`zcat $bed | wc -l`
echo "RH4-K27ac ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

bed="/home/gdstantonlab/mxw010/Data/Ben/pc-chip-seq/GSL-BS-2005/H3K27AC/RH30/align_primary/chip/2317294d-60c3-4998-9179-dc7da8c57608/call-call_peak/shard-0/execution/SEG0139_S7_L003_R1_001_val_1.srt.nodup_x_SEG0138_S6_L003_R1_001_val_1.srt.nodup.pval0.01.500K.narrowPeak.gz"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`zcat $bed | wc -l`
echo "RH30-K27ac ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

n1=`bedtools intersect -u -a FP-k27ac.bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a FP-k27ac.bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a FP-k27ac.bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a FP-k27ac.bed -b all_neg.bed | wc -l`
n5=`wc -l FP-k27ac.bed | awk '{print $1}'`
echo "FP-K27ac ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

bed="/home/gdstantonlab/mxw010/Data/Ben/pc-chip-seq/GSL-BS-2005/H3K27AC/RD/align_primary/chip/c28a7a00-d332-4997-bb07-59eedba9146c/call-call_peak/shard-0/execution/SEG0144_S2_L004_R1_001_val_1.nodup_x_SEG0143_S1_L004_R1_001_val_1.nodup.pval0.01.500K.narrowPeak.gz"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`zcat $bed | wc -l`
echo "RD-K27ac ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt


bed="/home/gdstantonlab/mxw010/Data/Ben/pc-chip-seq/GSL-BS-2005/H3K27AC/SMS-CTR/align_primary/chip/3b89bbf5-421e-444a-87c5-3bbe9b11dbff/call-call_peak/shard-0/execution/SEG0149_S7_L004_R1_001_val_1.srt.nodup_x_SEG0148_S6_L004_R1_001_val_1.srt.nodup.pval0.01.500K.narrowPeak.gz"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`zcat $bed | wc -l`
echo "CTR-K27ac ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

n1=`bedtools intersect -u -a FN-k27ac.bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a FN-k27ac.bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a FN-k27ac.bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a FN-k27ac.bed -b all_neg.bed | wc -l`
n5=`wc -l FN-k27ac.bed | awk '{print $1}'`
echo "FN-K27ac ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

#MYOD
bed="/home/gdstantonlab/mxw010/Data/Ben/MYOD/GSL-BS-2483/result/file_transfer/outputs/RH4/RH4_IDR_Conservative.bed"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`wc -l $bed | awk '{print $1}'`
echo "RH4-MYOD ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

bed="/home/gdstantonlab/mxw010/Data/Ben/MYOD/GSL-BS-2483/result/file_transfer/outputs/RH30/RH30_IDR_Conservative.bed"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`wc -l $bed | awk '{print $1}'`
echo "RH30-MYOD ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt


n1=`bedtools intersect -u -a FP-MYOD.bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a FP-MYOD.bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a FP-MYOD.bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a FP-MYOD.bed -b all_neg.bed | wc -l`
n5=`wc -l FP-MYOD.bed | awk '{print $1}'`
echo "FP-MYOD ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

bed="/home/gdstantonlab/mxw010/Data/Ben/MYOD/GSL-BS-2483/result/file_transfer/outputs/RD/RD_IDR_Conservative.bed"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`wc -l $bed | awk '{print $1}'`
echo "RD-MYOD ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

bed="/home/gdstantonlab/mxw010/Data/Ben/MYOD/GSL-BS-2483/result/file_transfer/outputs/SMS-CTR/SMS-CTR_IDR_Conservative.bed"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`wc -l $bed | awk '{print $1}'`
echo "CTR-MYOD ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt


n1=`bedtools intersect -u -a FN-MYOD.bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a FN-MYOD.bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a FN-MYOD.bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a FN-MYOD.bed -b all_neg.bed | wc -l`
n5=`wc -l FN-MYOD.bed | awk '{print $1}'`
echo "FN-MYOD ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

#P3F

bed="/home/gdstantonlab/mxw010/Data/Ben/downsampled/data/bed/RH4.bed"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`wc -l $bed | awk '{print $1}'`
echo "RH4-P3F ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

bed="/home/gdstantonlab/mxw010/Data/Ben/downsampled/data/bed/RH30.bed"
n1=`bedtools intersect -u -a $bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a $bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a $bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a $bed -b all_neg.bed | wc -l`
n5=`wc -l $bed | awk '{print $1}'`
echo "RH30-P3F ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt


n1=`bedtools intersect -u -a FP-P3F.bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a FP-P3F.bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a FP-P3F.bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a FP-P3F.bed -b all_neg.bed | wc -l`
n5=`wc -l FP-P3F.bed | awk '{print $1}'`
echo "FP-P3F ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

#MYCN
n1=`bedtools intersect -u -a MYCN_RH4.bed -b pos_to_neg.bed | wc -l`
n2=`bedtools intersect -u -a MYCN_RH4.bed -b neg_to_pos.bed | wc -l`
n3=`bedtools intersect -u -a MYCN_RH4.bed -b all_pos.bed | wc -l`
n4=`bedtools intersect -u -a MYCN_RH4.bed -b all_neg.bed | wc -l`
n5=`wc -l MYCN_RH4.bed | awk '{print $1}'`
echo "RH4-MYCN ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

#Total
n1=`wc -l pos_to_neg.bed | awk '{print $1}'`
n2=`wc -l neg_to_pos.bed | awk '{print $1}'`
n3=`wc -l all_pos.bed  | awk '{print $1}'`
n4=`wc -l all_neg.bed  | awk '{print $1}'`
n5=`wc -l /home/gdstantonlab/mxw010/Data/SPICE-C/RH4/Compartments/concordance/RH4_k27ac.cis.vecs.tsv | awk '{print $1}'`
echo "Total-all ${n1} ${n2} ${n3} ${n4} ${n5}" >> switch_table_tf.txt

sed -ie 's/-/ /g' switch_table_tf.txt



data <- read.table("switch_table_tf.txt", header = T)
data <- data[, -nrow(data)]

x <- data
x[, 3:7] <- t(apply(x[,-(1:2)], 1, function(y) y/y[length(y)]))
x <- x[-nrow(x), -ncol(x)]


x$cell <- factor(x$cell, levels = c("RH4", "RH30", "FP", "RD", "CTR", "FN"))
x$factor <- factor(x$factor, levels = c("K27ac", "MYOD", "P3F", "MYCN"))
df <- melt(x)

df$type <- factor(df$variable, levels=c('AB','BA','AA','BB'))


cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
ggplot(data=df, aes(x= type , y = value, color=factor, group=factor)) +
geom_line(size=1.5, linetype='twodash') +
facet_wrap(vars(cell)) +
scale_color_manual(values=cbPalette, name='Factor') +
ylab("Percentages")

x[x$cell == 'FP' | x$cell =='FN',]


y <- melt(x[x$cell == 'FP' | x$cell =='FN',])

#    cell factor         AB          BA        AA         BB
# 3    FP  K27ac 0.02638335 0.014475337 0.6323811 0.02583711
# 6    FN  K27ac 0.01060579 0.028862908 0.6370546 0.02474685
# 9    FP   MYOD 0.02571519 0.018090029 0.6030185 0.03901977
# 12   FN   MYOD 0.01441705 0.032107536 0.5698565 0.04819613
# 15   FP    P3F 0.04588235 0.009019608 0.5819608 0.04274510

#FP, P3F vs k27ac
subset(x, cell == 'FP')
prop.test(c(483,117),c(18307,2550))
#0.6796078 of consensus P3F sites in those 4 categories
#0.6990769 of consensus K27ac sites in those 4 categories
#FN, P3F vs k27ac
subset(data, cell == 'FN')

    subset(x, cell == 'FP')[,-(5:6)] %>%
    melt() %>%
    ggplot(aes(x= variable , y = value, fill=factor)) +
    geom_bar(position='dodge', stat='identity') +
    facet_wrap(vars(cell)) +
    scale_fill_manual(values=cbPalette, name='Factor') +
    ylab("Percentages")
```
# Fig1: Normalize Per Compartment/Bin

```{r per_bin}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
data <- read.table("switch_table.txt", header = T)
# data <- data[, -ncol(data)]
x <- data
x[, 3:6] <- sapply(3:6, function(i) data[, i] / data[nrow(data), i])
# x[nrow(x),] <- data[nrow(data),]
x$cell <- factor(x$cell, levels = c("RH4", "RH30", "FP", "RD", "CTR", "FN"))
x$factor <- factor(x$factor, levels = c("K27ac", "MYOD", "P3F", "MYCN"))
df <- melt(x[-nrow(x), -ncol(x)])

df$se <- unlist(sqrt((df$value * (1 - df$value)) / data[nrow(data), sapply(df$variable, match, table = colnames(data))]), use.names = F)
dodge <- position_dodge(width = 0.9)


subset(df, (cell == "FP" | cell == "FN")) %>%
    mutate(type = factor(variable, levels = c("AB", "BA", "AA", "BB"))) %>%
    ggplot(aes(x = type, y = value, fill = factor)) +
    # geom_line(size=1.5, linetype='twodash') +
    geom_bar(position = "dodge", stat = "identity", color = "black") +
    # geom_errorbar(aes(ymin=value-se, ymax=value+se), position=dodge, width=0.5, size=1.2) +
    facet_wrap(vars(cell)) +
    scale_fill_manual(values = cbPalette, name = "Factor") +
    ylab("Percentages")
```

# Fig2: Normalize by Number of binding sites (P3F has significantly fewer)

```{r per_binding}
data <- read.table("switch_table_tf.txt", header = T)
x <- data
x[, 3:7] <- t(apply(x[, -(1:2)], 1, function(y) y / y[length(y)]))
x$sites <- data$sites
x <- x[-nrow(x), ]
x$cell <- factor(x$cell, levels = c("RH4", "RH30", "FP", "RD", "CTR", "FN"))
x$factor <- factor(x$factor, levels = c("K27ac", "MYOD", "P3F", "MYCN"))
df <- melt(x[, -ncol(x)], id.vars = c("cell", "factor"))
total <- mapply(function(a, b) {
    data$sites[which(data$cell == a & data$factor == b)]
}, df$cell, df$factor)
df$se <- unlist(sqrt((df$value * (1 - df$value)) / total), use.names = F)
dodge <- position_dodge(width = 0.9)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
subset(df, (cell == "FP" | cell == "FN")) %>%
    mutate(type = factor(variable, levels = c("AB", "BA", "AA", "BB"))) %>%
    ggplot(aes(x = type, y = value, fill = factor)) +
    # geom_line(size=1.5, linetype='twodash') +
    geom_bar(position = "dodge", stat = "identity", color = "black") +
    #    geom_errorbar(aes(ymin = value - se, ymax = value + se), position = dodge, width = 0.5, size = 1.2) +
    facet_wrap(vars(cell)) +
    scale_fill_manual(values = cbPalette, name = "Factor") +
    ylab("Percentages")
```

# Fig3: Normalize by Per Bin Per 10000 Binding Sites
```{r}
data <- read.table("switch_table.txt", header = T)
# data <- data[, -ncol(data)]
x <- data
x[, 3:6] <- sapply(3:6, function(i) data[, i] / data[nrow(data), i])

y <- subset(x, cell == "FP" | cell == "FN")
y[, 3:6] <- y[, 3:6] / (y$sites / 10000)

ggplot(data = melt(y[, -ncol(y)]), aes(x = variable, y = value, fill = factor)) +
    # geom_line(size=1.5, linetype='twodash') +
    geom_bar(position = "dodge", stat = "identity", color = "black") +
    # geom_errorbar(aes(ymin=value-se, ymax=value+se), position=dodge, width=0.5, size=1.2) +
    facet_wrap(vars(cell)) +
    scale_fill_manual(values = cbPalette, name = "Factor") +
    ylab("Percentages")
```

# Fig4: Factor as Facets (compare against Fig1)

```{r facets}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
data <- read.table("switch_table.txt", header = T)
# data <- data[, -ncol(data)]
x <- data
x[, 3:6] <- sapply(3:6, function(i) data[, i] / data[nrow(data), i])
# x[nrow(x),] <- data[nrow(data),]
x$cell <- factor(x$cell, levels = c("RH4", "RH30", "FP", "RD", "CTR", "FN"))
x$factor <- factor(x$factor, levels = c("K27ac", "MYOD", "P3F", "MYCN"))
df <- melt(x[-nrow(x), -ncol(x)])

df$se <- unlist(sqrt((df$value * (1 - df$value)) / data[nrow(data), sapply(df$variable, match, table = colnames(data))]), use.names = F)
dodge <- position_dodge(width = 0.9)


subset(df, (cell == "FP" | cell == "FN")) %>%
    mutate(type = factor(variable, levels = c("AB", "BA", "AA", "BB"))) %>%
    ggplot(aes(x = type, y = value, fill = cell)) +
    # geom_line(size=1.5, linetype='twodash') +
    geom_bar(position = "dodge", stat = "identity", color = "black") +
    # geom_errorbar(aes(ymin=value-se, ymax=value+se), position=dodge, width=0.5, size=1.2) +
    facet_wrap(vars(factor)) +
    scale_fill_manual(values = cbPalette, name = "Factor") +
    ylab("Percentages")
```